\begin{abstract}
CARScenes is an annotation-only dataset and benchmark for structured scene understanding and failure analysis in autonomous driving. We canonicalize 5,192 front-camera frames from Argoverse1, Cityscapes, KITTI, and nuScenes into a versioned JSONL release with one schema-aligned record per frame. The release ships a machine-readable schema, validation and evaluation scripts, fixed split manifests, scenario-slice definitions, and Croissant metadata, while leaving upstream imagery under the original dataset licenses. The benchmark core covers environment, road geometry, traffic control, vehicles, pedestrians, visibility, ego context, and an ordinal severity label intended for stratified analysis rather than policy evaluation. Initial labels were produced with a VLM-assisted pipeline, then reviewed and corrected by a human annotator for every released record. CARScenes v1 freezes 4,592 training records, 500 silver development records for model selection, and a source-balanced 100-frame gold test split with a 25-frame agreement subset. The frozen release already exposes 1,586 traffic-light scenes, 1,781 scenes with vulnerable road users, and utility slices ranging from 28 rainy night-or-dusk scenes to 71 traffic-light scenes in the highest severity bucket. The current manuscript focuses on the reproducible release, the fully human-corrected benchmark protocol, and the scenario slices that make CARScenes useful for failure analysis across legacy driving datasets.
\end{abstract}
